\section*{Q2}

\subsection*{(a)}
Let $\B{OPT}(i,j)$ refer to the minimum cost of any alignment of $D[i]$ with $D[j]$

\subsection*{(b)}

Let $\B{COST}(a, b)$ refer to the cost of comparing characters $a$ and $b$. This returns $1$ if they are equal and $0$ if they are different. This is symmetric.

\begin{equation}
    \begin{split}
        \B{OPT}(i, j) = \begin{cases}
            0 & \dots \text{if } i = 0 \wedge j = 0 \\
            j & \dots \text{if } i = 0 \wedge j \neq 0\\
            i & \dots \text{if } j = 0 \wedge i \neq 0 \\
            \min\left(
            \begin{array}{l}
                \B{OPT}(i-1,j) + 1, \\
                \B{OPT}(i,j-1) + 1, \\
                \B{OPT}(i-1,j-1) + \B{COST}(D_1[i], D_2[j])
            \end{array}
            \right) & \dots \text{otherwise}
        \end{cases}
    \end{split}
\end{equation}

For initial conditions, we have base cases:
\begin{itemize}
    \item If $i = 0 \wedge j = 0$, then both strings are empty and both strings are identical, so any alignment has two identical strings which have 0 cost.
    \item If $i=0$, then $D_1[1 \dots i]$ is an empty string and $D_2[1 \dots j]$ is non empty. We align $D_1[1 \dots i]$ to match $D_2[1 \dots j]$ by inserting spaces, but $D_2[j] \in \{'a', 'c', 't', 'g'\}$. So all characters in $D_2[1 \dots j]$ mismatch and this is true for any alignment, so the minimum cost is the length of $D_2[j]$, $j$.
    \item If $j=0$, then $D_2[j]$ is an empty string and $D_1[j]$ is non empty. The same argument made for $i=0$ applies here.
\end{itemize}

For the last case, there are three sub cases to consider:
Suppose 
\begin{enumerate}
    \item Add space for $D_1$\\
    Let $A$ and $B$ be any alignment corresponding to $\B{OPT}(i,j-1)$. We can generate a new alignment for $D_1[i]$ and $D_2[j]$ by appending $D_2[j] \in \{'a', 'c', 't', 'g'\}$ to $B$ and a space to $A$. They are necessarily different, so this adds a cost of 1.
    \item Add space for $D_2$\\
    Let $A$ and $B$ be any alignment corresponding to $\B{OPT}(i-1,j)$. We can generate a new alignment for $D_1[i]$ and $D_2[j]$ by appending $D_1[i] \in \{'a', 'c', 't', 'g'\}$ to $A$ and a space to $B$. They are necessarily different, so this adds a cost of 1.
    \item Accept a match or mismatch\\
    Let $A$ and $B$ be any alignment corresponding to $\B{OPT}(i-1,j)$. We can generate a new alignment for $D_1[i]$ and $D_2[j]$ by appending $D_1[i]$ to $A$ and appending $D_2[j]$ to B. The additional cost is $0$ if they match or $1$ if they mismatch.
\end{enumerate}

\subsection*{(c)}
We try a bottom up approach.

\begin{verbatim}
C = []
for i in 1..n:
    for j in j..1:
        C[i,j] = OPT(i,j)
\end{verbatim}

This guarantees that $\B{OPT}(i-1,j)$, $\B{OPT}(i,j-1)$ and $\B{OPT}(i-1,j-1)$ have all been computed before calculating $\B{OPT}(i,j)$

\subsection*{(d)}
\paragraph{Space Complexity:}
We need to store a cache table of size $n_1+1 by n_2+1$, which would be $O(n_1n_2)$. Notice that we only need $\B{OPT}(i-1,j)$, $\B{OPT}(i,j-1)$ and $\B{OPT}(i-1,j-1)$, so we can ignore all columns before $i-1$, so the actual space requirement is closer to $O(min(n_1,n_2)$ If you swap the two so that the number of columns is smaller than the number of rows.

\paragraph{Time Complexity:}
For each slot in our cache grid, calculating the optimal value only requires considering a constant amount of work. Since there are $O(n_1n_2)$ entries, the running time is $O(n_1n_2)$.

\subsection*{(e)}

\begin{verbatim}
# Initial conditions for backwards trace
S_1 = []
S_2 = []
i = n_1
j = n_2

# Function for the cost function stated above
def Cost(a,b):
    if a == b:
        return 1
    else:
        return 0

while true:
    if i > 0 and j > 0 and C[i][j] == C[i-1][j-1] + Cost(D_1[i-1],D_2[j-1]):
        # Safely check if we chose OPT(i-1,j-1)
        S_1 += D_1[i-1]
        S_2 += D_2[i-1]
        i -= 1
        j -= 1
    elif i > 0 and C[i][j] == C[i-1][j]:
        # Safely check if we chose OPT(i-1,j)
        S_1 += D_1[i-1]
        S_2 += " "
    elif j > 0 and C[i][j] == C[i][j-1]:
        # Safely check if we chose OPT(i,j-1)
        S_1 += " "
        S_2 += D_2[j-1]
    elif i == 0 and j == 0:
        # We've fully traversed back
        break

# Since we built the strings backwards, we need to reverse them.
S_1.reverse()
S_2.reverse()
\end{verbatim}